Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2023-03-05 latex/comp_or.tex
Linear relaxation

We converted Shims into a process-based parallel computing heuristic that quickly finds good solutions for a wide range of problem sizes, an essential contribution as it surpassed all Shims results and some of the integer programming results, all methods under a time limit to guarantee that a distribution plan is yielded in less than one hour.

@@

\bgroup
\def\arraystretch{1.2}
\begin{table}[H]
	\centering
	\small
	\begin{tabular}{MT}
		&Informally, ACLP+RPDP can be summarized as follows:\\
		\midrule
		max &  (items score sum) / (tour cost) of picked up and delivered items on a tour  \\
		\midrule
s.t.    & Along a tour, the set of unvisited nodes is updated. \\
		& In each node, an item may be included in at most one pallet.\\		
		& In each node, consolidated items are composed of items with the same destination. \\
		& Weight, volume and score of a consolidated item are the corresponding sum of their components.\\
		& Consolidated items remain on board until their destinations.\\	
		& Consolidated items can only be included in the same pallet if their destinations are the same.\\
		& Only items destined for the remaining nodes can be loaded.  \\
		& The lateral and longitudinal torques must be within the operational range of the aircraft.\\
		& Weight and volume limitations of pallets must be respected.\\
		& The total weight must be less than the aircraft payload or the total pallet capacity, whichever is the lowest.\\	
		& The pallets destined for the next node of the tour should be put as near as possible to the cargo ramp door (our inclusion).\\		
		\midrule
	\end{tabular}
	\normalsize
\end{table}
\egroup

@@

		\State $\pi(0) \gets 0$ \label{tour:pi1}
		\State $\pi(K+1) \gets 0$ \label{tour:pi2}
		\State $score \gets 0$ \label{tour:score}
		\State $cost \gets 0$ \label{tour:cost}
		\For {$k\gets0$ to $K$} \label{tour:loop1}		
			\For{$i \gets 1$ to $m$}
				\State $p_i.to[k] \gets -1$ \label{tour:-1} \Comment{reset this pallet destination}
			\EndFor	
			\If {$k = 0$}
				\State $L_0 \gets L$
				\State Let $G_1(M \cup N_0, \varnothing)$ \label{tour:g11}
			\Else
				\State $L_k \gets L_k - \pi(k)$  \label{tour:lk1}			
				\State $Q_{\pi(k)}, E^Q_{\pi(k)}, M \gets UpdateConsolidated(\pi(k))$ \label{tour:dest}			
				\State Let $G_1(M \cup N_{\pi(k)} \cup Q_{\pi(k)}, E^Q_{\pi(k)})$ \label{tour:g12}
			\EndIf  \label{tour:lk2}	
			\State $M \gets SetPalletsDestination( \pi(k) )$ \label{tour:dest2}		
			\State $G_2 \gets SolveNode(method, \pi(k), G_1, timeLim)$ \label{tour:node}
			\State $s, \epsilon \gets ScoreAndDeviation(\pi(k), G_2)$ \label{tour:analyse}
			\State $score \gets score + s$ \label{tour:score2}
			\State $cost \gets cost + c_{\pi(k),\pi(k+1)} * (1 + c_g * |\epsilon|)$ \label{tour:cost2} 
		\EndFor  \label{tour:loop2}
		\Return $score / cost$ \label{tour:f}

@@

@@

		{\it Stopping criterion:} & All methods have the same time limit of $1.8$ seconds per node.\\

@@

\multirow{7}{*}{$1.2$}&\multirow{2}{*}{ $Gurobi$}& $f$                & {\bf 12.00} & {\bf 8.38}  & {\bf 12.19} & {\bf 13.38} & {\bf 14.42} & 13.25        & {\bf 0.99} \\% 7412
&                                                & {\it run-time (s)} &  8          &   14        &  25         &  95         &    217      &  1719        &  1.0       \\%2078

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                       & $f$                & 9.69        &  6.37       &  9.58       &  11.07      &   12.24     & 11.94        &  0.82\\%
&                                                & {\it run-time (s)} &  < 1        &  3          &    5        &   11        &   27        & 342          &  5.3\\%389

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                     & $f$                & 11.90       &  7.61       &  10.86      & 13.13       & 13.75       & {\bf 13.75}  &  0.96\\%
&                                                & {\it run-time (s)} &  < 1        &  < 1        &    3        &  9          &  26         & 156          &  {\bf 10.6}\\%

@@

\multirow{7}{*}{$1.5$}&\multirow{2}{*}{ $Gurobi$}& $f$                & 16.49       & {\bf 11.60} & {\bf 16.68} & {\bf 18.33} & {\bf 20.10} & 17.94        & {\bf 0.99}\\%10209
&                                                & {\it run-time (s)} &  12         &   25        &  45         &  169        &   437       &  2377        & 1.0\\%3065

@@

\midrule
\multirow{7}{*}{$2.0$}&\multirow{2}{*}{ $Gurobi$}& $f$                & {\bf 24.04} & {\bf 17.33} & {\bf 24.62} & {\bf 26.36} & {\bf 28.88} & {\bf 25.89}  & {\bf 1.00}\\%14712
&                                                & {\it run-time (s)} &  14         &   34        &  62         &  236        &  609        & 3431         & 1.0\\%4836

@@

For the MIP solver in some scenarios, less than 50\% of the results were optimal, and the remaining were suboptimal due to the time limit. This explains that {\it mpShims} surpassed some of the MIP solver results.

In \cite{MesquitaSanches2023}, {\it Shims} surpassed the MIP solver, on average, because the time limit they used (0.7s) was sufficient to guarantee a worst-case solution in less than 30min. This decision caused most of the MIP solver results to be poor, making {\it Shims} outstanding as it was the unique method to perform well with all scenarios and instances.

In this work, we used as a time limit the worst {\it Shims} run-time, which was 1.8s. This made the MIP solver yield optimal or near-optimal solutions. This is good for the reader to have an idea of the real performance differences among the methods.

On average, {\it mpShims} is 14 times faster than the MIP solver, and it reached 96\% of {\it Gurobi} $f$ values, on average, which is good for a tool of operational use. The use of parallelism is another important contribution of this work. It has improved {\it Shims} results and also its speed.

\subsection{Yet another improvement}

@@

Let $Z^{\pi(k)}_{iq}$ be the set of decision variables that relates pallet $i$ to consolidated $q$ in node $\pi(k)$.

@@

@@

>>>>

We converted Shims into a process-based parallel computing heuristic that quickly finds good solutions for a wide range of problem sizes, an essential contribution as it surpassed all Shims results.

@@

\bgroup
\def\arraystretch{1.2}
\begin{table}[H]
	\centering
	\small
	\begin{tabular}{MT}
		&Informally, ACLP+RPDP can be summarized as follows:\\
		\midrule
		max &  (items score sum) / (tour cost) of picked up and delivered items on a tour  \\
		\midrule
s.t.    & Along a tour, the set of unvisited nodes is updated. \\
		& In each node, an item may be included in at most one pallet.\\		
		& In each node, consolidated items are composed of items with the same destination. \\
		& Weight, volume and score of a consolidated item are the corresponding sum of their components.\\
		& Consolidated items remain on board until their destinations.\\	
		& Consolidated items can only be included in the same pallet if their destinations are the same.\\
		& Only items destined for the remaining nodes can be loaded.  \\
		& The lateral and longitudinal torques must be within the operational range of the aircraft.\\
		& Weight and volume limitations of pallets must be respected.\\
		& The total weight must be less than the aircraft payload or the total pallet capacity, whichever is the lowest.\\	
		& {\bf The pallets destined for the next node of the tour should be put as near as possible to the cargo ramp door (our inclusion)}.\\		
		\midrule
	\end{tabular}
	\normalsize
\end{table}
\egroup

@@

		\State $\pi(0) \gets 0$ \label{tour:pi1}
		\State $\pi(K+1) \gets 0$ \label{tour:pi2}
		\State $score \gets 0$ \label{tour:score}
		\State $cost \gets 0$ \label{tour:cost}
		\For {$k\gets0$ to $K$} \label{tour:loop1}		
			\For{$i \gets 1$ to $m$}
				\State $p_i.to[k] \gets -1$ \label{tour:-1} \Comment{reset this pallet destination}
			\EndFor	
			\If {$k = 0$}
				\State $L_0 \gets L$
				\State Let $G_1(M \cup N_0, \varnothing)$ \label{tour:g11}
			\Else
				\State $L_k \gets L_k - \pi(k)$  \label{tour:lk1}			
				\State $Q_{\pi(k)}, E^Q_{\pi(k)}, M \gets UpdateConsolidated(\pi(k))$ \label{tour:dest}			
				\State Let $G_1(M \cup N_{\pi(k)} \cup Q_{\pi(k)}, E^Q_{\pi(k)})$ \label{tour:g12}
			\EndIf  \label{tour:lk2}	
			\State $M \gets SetPalletsDestination( \pi(k) )$ \label{tour:dest2}		
			\State $G_2 \gets SolveNode(method, \pi(k), G_1, timeLim)$ \label{tour:node}

{\color{blue} \State	$G_3 \gets minRampDist(G_2)$  \label{tour:minRampDist} }

			\State $s, \epsilon \gets ScoreAndDeviation(\pi(k), G_3)$ \label{tour:analyse}
			\State $score \gets score + s$ \label{tour:score2}
			\State $cost \gets cost + c_{\pi(k),\pi(k+1)} * (1 + c_g * |\epsilon|)$ \label{tour:cost2} 
		\EndFor  \label{tour:loop2}
		\Return $score / cost$ \label{tour:f}

@@

In line \ref{tour:minRampDist}, $minRampDist$ minimizes the distance of the next node destined pallets, keeping the CG in its operational range. This precedure is implemented in integer programming as presented in Subsection \ref{improvement}, according to equations \ref{eq:distobj} to \ref{eq:disttorque}.

@@

@@

\multirow{7}{*}{$1.2$}&\multirow{1}{*}{ $Gurobi$}& $f$                &{\bf 11.22}   &{\bf 7.88}   & 10.25      & 11.02       & 11.88        & 12.92       & 0.90 \\% 6484    7222

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                       & $f$                & 9.66         &  6.38       &  9.55       &  11.12      &  12.37      &   12.03      & 0.85 \\%6111
&                                                & {\it run-time (s)} &  < 1         &  5          &    11       &   43        &  128        &  835        &  1.0 \\%1023

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                     & $f$                & 11.03        & 7.43        &{\bf 12.58}  &{\bf 13.08}  &{\bf 14.02}  & {\bf 13.44}  &  {\bf 0.99}\\%7158

&                                                & {\it run-time (s)} &  < 1         &  2          &    4        &  15         &  47         & 295          &  {\bf 2.8}\\%364

@@

\multirow{7}{*}{$1.2$}&\multirow{1}{*}{ $Gurobi$}& $f$                &{\it 0}   &{\it 0}   &{\it 0}  &{\it 0}  &{\it 0}   &{\it 0}  & {\it 1.00} \\% 6607

@@

\midrule
\multirow{7}{*}{$1.2$}&\multirow{1}{*}{ $Gurobi$}& $f$                &{\it 0}   &{\it 0}   &{\it 0}  &{\it 0}  &{\it 0}   &{\it 0}  & {\it 1.00} \\% 6607

@@

The upper bound is the objective function value of an LP relaxation of the problem, proving that no integer feasible solution of larger objective value than this value can exist.

In an attempt to find the upper bound for $f$, we solved all scenarios and instances with {\it Gurobi} in a linear relaxed model. We applied a time limit of 60s in the expectation that it would be enough for {\it Gurobi} to determine the upper bounds. This is good for the reader to have an idea of the real performance differences between serial and parallel {\it Shims}.

Most of the {\it mpShims} $f$ results are superior to the upper bound found because they are not really the upper bound. As {\it Gurobi} runtime was limited, it did not have enough time to update the upper bound. But it is a good reference to compare {\it Shims} and {\it mpShims} performances.

As to the speedup, on average, {\it mpShims} was 2.8 times faster than {\it Shims}. Theoretically, it could be more, but due to the burden of sharing the CG deviation and the list of items among the processes for reading and writing, the runtime management to hinder race condition limited {\it mpShims} performance.

\subsection{Yet another improvement}
\label{improvement}

@@

Let $Z^{\pi(k)}_{iq}$ be the set of binary decision variables that relates pallet $i$ to consolidated $q$ in node $\pi(k)$.

@@

% \ref{eq:distobj} to \ref{eq:disttorque}

@@

We added an important contribution by favoring ground operations reducing in 29\%, in average, the distances pallets will be moved on disembarking operations.
